---
title: "BS9001 Research Experience: Bootstrap Preparation and Resampling Protocol"
author: "Austin Chia Cheng En (U1740366A)"
output:
    rmdformats::material:
      highlight: kate
      code_folding: "hide"
---

# 1) Importing necessary libraries

##### All the necessary libraries are first imported. `readr` for writing dataframes into csv files and `genefilter` for the `rowttests` function.

```{r Calling Libraries, eval=FALSE}
library(readr)
library("genefilter")
```

# 2) Split dataframe into 2 classes - disease and control, taken from metadata

###### The metadata of the GDS was accessed and columns 1 to 12 are for the control group and columns 38 to 49 are for the disease class (schizophrenia). The expression data of both groups were subsetted.

```{r 2) Split dataframe into 2 classes - disease and control, taken from metadata, eval=FALSE}
data <- read.csv("GDS3345_mean_df.csv", header = TRUE, stringsAsFactors =  FALSE, sep = ",")
rownames(data) <- rownames(mean_df)
data <- cbind(data[,1:12], data[,38:49])

# control class - 1-12
# disease class - 13-24

```

# 3) Appending binary significance t-test results to matrix, sampled 1000 times original loop

##### A progress bar was first added through the "progress" package.

##### A factor with the corresponding "normal" and "disease" classes was defined.

##### Working input dataframe was coerced into a matrix for the rowttest function.

##### An empty list was initialized to store output data from the loop.

```{r 3) Appending binary significance t-test results to matrix, sampled 1000 times original loop, eval=FALSE}

# Genefilter package method
# BiocManager::install("genefilter")

library(progress)
pb <- progress_bar$new(total = 1000)
fac <- c(rep("normal",4),rep("disease",4))
data_mat <- as.matrix(data)
boot_list <- list()

for (i in 1:1000) 
{
  my_significant_genes <- c()
  control <- as.matrix(sample(data[,1:12], size = 4, replace = TRUE))
  disease <- as.matrix(sample(data[,13:24], size = 4, replace =TRUE))
  m3 <- cbind(control, disease)
  test_ttest <- rowttests(m3,fac = as.factor(fac))
  my_significant_genes_v2 <- as.numeric(test_ttest$p.value < 0.05)
  boot_list <- append(boot_list, list(my_significant_genes_v2))
  
  pb$tick()
  Sys.sleep(1 / 1000)
}

```

# 4) Transforming nested list into dataframe

###### The nested list output derived from the `for` loop was coerced into a dataframe.

```{r 4) Transforming nested list into dataframe, eval=FALSE}

boot_df <- data.frame(matrix(unlist(boot_list), nrow=nrow(data), byrow=F),stringsAsFactors=FALSE)

```

# 5) Jaccard coefficient


##### A Jaccard coefficient function was defined. A progress bar was included to view run progress.

```{r 5) Defining a Jaccard coefficient function, eval=FALSE}
jac_func <- function (x, y) 
{
  M_11 <- sum(x == 1 , y == 1)
  M_10 <- sum(x == 1 , y == 0)
  M_01 <- sum(x == 0 , y == 1)
  return (M_11 / (M_11 + M_10 + M_01))
}

jac_df <- data.frame(matrix(data = NA, nrow = length(boot_df), ncol = length(boot_df)))

pb <- progress_bar$new(total = 1000)

```

##### The Jaccard coefficients were assigned into a table through the use of a nested for loop. The bottom half of the table is not filled up, as the values will be a mirror image of those at the top half.
```{r Assigning Jaccard coefficients into a table, eval=FALSE}
for (r in 1:length(boot_df)) 
  {
  for (c in 1:length(boot_df)) 
    {
    if (c == r) {
      jac_df[r,c] = 1
    } else if (c > r) {
      jac_df[r,c] = jac_func(boot_df[,r], boot_df[,c])
    }
  }
  pb$tick()
  Sys.sleep(1 / 1000)
}

variable_names <- sapply(boot_df, attr, "label")
colnames(jac_df) <- paste0("S", seq(1:ncol(jac_df)))
rownames(jac_df) <- paste0("S", seq(1:nrow(jac_df)))
jac_mat <- matrix(jac_df)

```



